---
title: 'This is a note for Chen Lab on GSA analysis'
author: 'Jianying Li'
date: '13 February 2020'
output:
  html_document:
    number_sections: true
    toc: true
    fig_width: 7
    fig_height: 4.5
    theme: readable
    highlight: tango
---

# Introduction

Yicheng helped David in Dr. Chen's lab to write a SAM-GSA code for microarray analysis. Here is a note for such analysis.

Here are the files that are needed for the analysis

* Normalized microarray data
* GSA geneset
* This program.

# Load and check data

```{r, message = FALSE}
# Load packages
library(samr)
library(GSA)
```

Now that our packages are loaded, let's read in and take a peek at the data.

Microarray Data 
```{r, message=FALSE, warning=FALSE}

setwd("X:/project2020/MustARD/GSA_analysis/htmlReport/")


load ("X:/project2019/RNAseqProj/results/firstPassAnalysis/K70ALL4_DEseq2.rda")
normalizedReadCountsAll = deseq2.results.K70.all4$norm.dat
p10.dt <- normalizedReadCountsAll[,c(1:3, 7,8)]
sample.names <- colnames(p10.dt )
gene.names <- rownames(p10.dt )
x1 = as.matrix (p10.dt)
str(matrix(rnorm(1000*9),ncol=5))
y1 <- c(rep(1, 3), rep(2, 2))
samfit1 <- SAM(as.matrix(x1), y1, resp.type = "Two class unpaired")
plot(samfit1)
```


# Running SAM analysis

```{r, message=FALSE, warning=FALSE}
#Getting the class label

MD.dt <- normalizedReadCountsAll[,c(1:6)]
sample.names <- colnames(MD.dt )
gene.names <- rownames(MD.dt )
x2 = as.matrix (MD.dt)
str(matrix(rnorm(1000*9),ncol=6))
y2 <- c(rep(1, 3), rep(2, 3))
samfit2 <- SAM(as.matrix(x2), y2, resp.type = "Two class unpaired")
plot(samfit2)
```

Fit the SAM analysis model and plot the results for P10

```{r results='asis'}
samfit1 <- SAM(as.matrix(x1), y1, resp.type = "Two class unpaired")
plot(samfit1)
```


Fit the SAM analysis model and plot the results for MD

```{r results='asis'}
samfit1 <- SAM(as.matrix(x2), y2, resp.type = "Two class unpaired")
plot(samfit1)
```

Prepare files for GSA now

```{r, message=FALSE, warning=FALSE}

x1 = x2
y1 = y2

gene.index <- paste ("g", c(1:dim(x1)[1]), sep="")
z <- cbind (as.data.frame (list ("X"= gene.index, "GeneSymbol"= gene.names)), x1)


colnames(z)[1:2] <- c("X", "GeneSymbol")
z <- as.data.frame(z)
#Make "Gene name" as string
z$X <- as.character(z$X)
#Make "Gene symbol"" as string
z$GeneSymbol <- as.character(z$GeneSymbol)
data <- list(x = x1, y = y1, geneid = z$X, genenames = z$GeneSymbol , logged2 = TRUE)


```

# Working on GSA
## Basic set up

```{r , message=FALSE, warning=FALSE}

## Setting the random seed for future validation
set.seed(1234567)

```


## Load gene set database


```{r , message=FALSE, warning=FALSE}


clean.gmt.data <- function (db)
{
 temp.geneset <- c(list())
 for (i in 1:length(db$genesets))
 {
   temp.geneset[[i]] <- db$genesets[[i]]
   if (length(which(temp.geneset[[i]]=="")) > 0 )
   {
      temp.geneset[[i]]  <- temp.geneset[[i]][-which(temp.geneset[[i]]=="")]
   }
 }
 db$genesets <- temp.geneset 
 return (db)
}



# original file
#c1 <- GSA.read.gmt("X:/project2020/MustARD/GSA_analysis/GSAdb/Knowledge_based_gene_sets_20200213.gmt")
c1 <- GSA.read.gmt("X:/project2020/MustARD/GSA_analysis/GSAdb/ESCC_collections_20200213.gmt")

c1.cleaned <- clean.gmt.data(c1)
str(c1)
str(c1$genesets)

print ("Gene sets in database file (20200213).gmt")
c1$geneset.names

```

## Run GSA


```{r  results = 'hide', message=FALSE, warning=FALSE}
c1 <- c1.cleaned

gsafit <- GSA(as.matrix(x1), y1, c1$genesets, z$GeneSymbol, resp.type = "Two class unpaired", nperms = 1000)

## Do I need the unique gene symbol??
#gsafit <- GSA(x, y, c1$genesets, unique(z$GeneSymbol), resp.type = "Two class unpaired", nperms = 1000)


str(gsafit)
plot(gsafit$GSA.scores)

```



```{r ,  message=FALSE, warning=FALSE}
gsafit$fdr.lo
gsafit$fdr.hi

# To many and will NOT print it to the html file
# gsafit$gene.scores

GSA.plot(gsafit)

GSA.listsets(gsafit, c1$geneset.names, FDRcut=.5)
GSA.correlate(c1$genesets, z$GeneSymbol)

gsa2 <- GSA.func(as.matrix(x1), y1, c1$genesets, z$GeneSymbol, resp.type = "Two class unpaired")
#gsa2 <- GSA.func(x, y, c1$genesets, unique(z$GeneSymbol), resp.type = "Two class unpaired")
str(gsa2)



```

## More on GSA


```{r , message=FALSE, warning=FALSE}



#for loop from 1:length(gsals$negative[,1])
GSA.genescores(5, c1$genesets, gsafit, z$GeneSymbol)
GSA.plot(gsafit, fac = 1, FDRcut = 1)

gsals <- GSA.listsets(gsafit, c1$geneset.names, FDRcut=.5)
str(gsals)
str(gsals$negative)
gsals$negative[,1]
length(gsals$negative[,1])
as.numeric(gsals$negative[,1][1])
GSA.genescores(as.numeric(gsals$negative[,1][1]), c1$genesets, gsafit, z$GeneSymbol)


```


## Export results


```{r , message=FALSE, warning=FALSE}
## sink() function does not work well with rmarkdown, 

#sink("V1-GSA-TF-negative.txt")
for (i in 1:length(gsals$negative[,1])){
  print(gsals$negative[,2][i])
  print(GSA.genescores(as.numeric(gsals$negative[,1][i]),
                 c1$genesets, gsafit, z$GeneSymbol))
  cat("\n")
  
}
#sink()

#sink("V1-GSA-TF-positive.txt")

for (i in 1:length(gsals$positive[,1])){
  print(gsals$positive[,2][i])
  print(GSA.genescores(as.numeric(gsals$positive[,1][i]),
                       c1$genesets, gsafit, z$GeneSymbol))
  cat("\n")
  
}
#sink()

#print GSA -TF output
#sink("V1-GSA-TF-output.txt")
print(GSA.listsets(gsafit, c1$geneset.names, FDRcut=.5))
#sink()

#Print negative names

#sink("v2-negativenames.txt")
print ("Negaivie names")
for (i in as.numeric(gsals$negative[,1])){
  print(c1$geneset.names[i])
  
}
#sink()

#print positivenames

#sink("V1-positivenames.txt")
print ("Positivie names")
for (i in as.numeric(gsals$positive[,1])){
  print(c1$geneset.names[i])
  
}
#sink()

```


# Conclusion

This is a valid way to document analysis.

---